---
title: "expercise2"
author: "Lukas Bieri (bieriluk)"
date: "05.01.2023"
format: html
editor: source
---
# Exercise 2 -  Computational Movement Analysis: Patterns and Trends in Environmental Data

## 1. Information & Metadata

### 1.1 Information

**Study program:** MSc in Environment and Natural Resources FS23

**Module:** Computational Movement Analysis: Patterns and Trends in Environmental Data

**Course:** R Exercise 2 - Week 2

**Lecturers:** Patrick Laube (laup), Nils Ratnaweera (rata) & Dominic Lüönd (luoe)

**Authors:** Lukas Bieri (bieriluk)

### 1.1 Structure R-Project

The following project was created for the exercise:
cma-week2-rexercise

Data was stored in the folder "datasets".

Code and explanations in this Quarto-File where stored and submitted via the public GitHub Repo "bieriluk/cma-week2":
https://github.com/bieriluk/cma-week2.git

### 1.2 Data

Data for the exercise was provided by the teaching staff via the plattform "Moodle". 

The data originates from the ZHAW Research Projekt "Using an acoustic signal to prevent wild boars from damaging crops"
https://www.zhaw.ch/en/about-us/news/news-releases/news-detail/event-news/using-an-acoustic-signal-to-prevent-wild-boars-from-damaging-crops/

The following datasets were used:

1.  File: "wildschwein_BE_2056.csv"
    Downloaded: 28.04.2023
    Source: https://moodle.zhaw.ch/pluginfile.php/1168373/mod_folder/content/0/wildschwein_BE_2056.csv?forcedownload=1
    
Information on the R Exercise can be found here:
https://computationalmovementanalysis.github.io/FS23/

These R Exercises are created by Patrick Laube, Nils Ratnaweera, Nikolaos Bakogiannis and Dominic Lüönd for the Course Computational Movement Analysis and are licensed under Creative Commons Attribution 4.0 International License.

### 1.4 Used Software

**R version 4.2.1 (2022-06-23 ucrt)** -- "Funny-Looking Kid"
Copyright (C) 2022 The R Foundation for Statistical Computing
Platform: x86_64-w64-mingw32/x64 (64-bit)

**RStudio 2023.03.0+386** "Cherry Blossom" Release (3c53477afb13ab959aeb5b34df1f10c237b256c3, 2023-03-09) for Windows
Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) RStudio/2023.03.0+386 Chrome/108.0.5359.179 Electron/22.0.3 Safari/537.36

**Gitfor WindowsVersion 2.40.1**


{{< pagebreak >}}

## 2. Preparation

In Preparation: Restart R and clear console.

Then: Load necessary functions

```{r}
library("readr") # to import tabular data (e.g. csv)
library("dplyr") # to manipulate (tabular) data
library("ggplot2") # to visualize data
library("sf")
```


## Import wild boar data & convert with to spatial data
```{r}

wildschwein_BE <- read_delim("datasets/wildschwein_BE_2056.csv", ",")

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)

wildschwein_BE
```
We need N & E later, so add remove = FALSE to not remove them wehn adding geometry


## How many individuals were tracked 
```{r}
wildschwein_BE$TierName |> unique()
```


## For how long were the individual tracked? Are there gaps? Were all individuals tracked concurrently or sequentially?
```{r}
ggplot(wildschwein_BE, aes(DatetimeUTC, TierName)) +
  geom_point()
```


## What is the temporal sampling interval between the locations?
```{r}
wildschwein_BE <- group_by(wildschwein_BE, TierID)

wildschwein_BE <- mutate(wildschwein_BE, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

wildschwein_BE

ggplot(wildschwein_BE, aes(timelag/60)) +
  geom_histogram(binwidth = 1) +
  lims(x = c(0, 5000/60)) +
  scale_y_log10()

wildschwein_BE |> 
  filter(DatetimeUTC < "2014-08-24") |> 
  ggplot(aes(DatetimeUTC, timelag, colour = TierName)) +
  geom_point() + 
  geom_line()
```

--> Sampling rate different during day and night

## Task 3 Deriving speed

Euclidian Distance (use dummy values to calculate distance first before using all data)
```{r}
N1 <- 1204752
N2 <- 1204863
E1 <- 2570409
E2 <- 2570402

sqrt((E1-E2)^2 + (N1-N2)^2)

```
--> Formula works, implement it in the data frame

```{r}
wildschwein_BE <- wildschwein_BE |> 
  group_by(TierName) |> 
  mutate(steplength_m = sqrt((E-(lead(E))^2 + (N-lead(N))^2)))

wildschwein_BE <- wildschwein_BE |> 
  mutate(speed_ms = steplenth_m/timelag)
```

