---
title: "expercise2"
author: "Lukas Bieri (bieriluk)"
date: "05.01.2023"
format: html
editor: source
---
# Exercise 2 -  Computational Movement Analysis: Patterns and Trends in Environmental Data

## 1. Information & Metadata

### 1.1 Ordnerstruktur R-Projekt

The following project was created for the exercise:
cma-week2-rexercise

Data was stored in the folder "datasets".

### 1.3 Data

Data for the exercise was provided by the teaching staff in the module an provided via the plattform "Moodle". 

The following datasets were used:

1.  File: "wildschwein_BE_2056.csv"
    Downloaded: 28.04.2023
    Source: https://moodle.zhaw.ch/pluginfile.php/1168373/mod_folder/content/0/wildschwein_BE_2056.csv?forcedownload=1


### 1.4 Used Software

**R version 4.2.1 (2022-06-23 ucrt)** -- "Funny-Looking Kid"
Copyright (C) 2022 The R Foundation for Statistical Computing
Platform: x86_64-w64-mingw32/x64 (64-bit)

**RStudio 2023.03.0+386** "Cherry Blossom" Release (3c53477afb13ab959aeb5b34df1f10c237b256c3, 2023-03-09) for Windows
Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) RStudio/2023.03.0+386 Chrome/108.0.5359.179 Electron/22.0.3 Safari/537.36

**Gitfor WindowsVersion 2.40.1**


{{< pagebreak >}}

## 2. Preparation

In Preparation: Restart R and clear console.

Then: Load necessary functions

```{r}
library("readr") # to import tabular data (e.g. csv)
library("dplyr") # to manipulate (tabular) data
library("ggplot2") # to visualize data
library("sf")
```


## Import wild boar data & convert with to spatial data
```{r}

wildschwein_BE <- read_delim("datasets/wildschwein_BE_2056.csv", ",")

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)

wildschwein_BE
```
We need N & E later, so add remove = FALSE to not remove them wehn adding geometry


## How many individuals were tracked 
```{r}
wildschwein_BE$TierName |> unique()
```


## For how long were the individual tracked? Are there gaps? Were all individuals tracked concurrently or sequentially?
```{r}
ggplot(wildschwein_BE, aes(DatetimeUTC, TierName)) +
  geom_point()
```


## What is the temporal sampling interval between the locations?
```{r}
wildschwein_BE <- group_by(wildschwein_BE, TierID)

wildschwein_BE <- mutate(wildschwein_BE, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

wildschwein_BE

ggplot(wildschwein_BE, aes(timelag/60)) +
  geom_histogram(binwidth = 1) +
  lims(x = c(0, 5000/60)) +
  scale_y_log10()

wildschwein_BE |> 
  filter(DatetimeUTC < "2014-08-24") |> 
  ggplot(aes(DatetimeUTC, timelag, colour = TierName)) +
  geom_point() + 
  geom_line()
```

--> Sampling rate different during day and night

## Task 3 Deriving speed

Euclidian Distance (use dummy values to calculate distance first before using all data)
```{r}
N1 <- 1204752
N2 <- 1204863
E1 <- 2570409
E2 <- 2570402

sqrt((E1-E2)^2 + (N1-N2)^2)

```
--> Formula works, implement it in the data frame

```{r}
wildschwein_BE <- wildschwein_BE |> 
  group_by(TierName) |> 
  mutate(steplength_m = sqrt((E-(lead(E))^2 + (N-lead(N))^2)))

wildschwein_BE <- wildschwein_BE |> 
  mutate(speed_ms = steplenth_m/timelag)
```

